---
title: "Evaluation of viral load trajectory simulation runs"
output: html_document
date: "2025-11-25"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(stringr)
library(ggplot2)
library(RColorBrewer) # For color palettes
library(xtable)
library(here)
library(tools)


TOP <- here()
source(file.path(TOP, "sars2vl", "data_params.R"))
source(file.path(TOP, "sars2vl", "plot_params.R"))
source(file.path(TOP, "sars2vl", "plot_utils.R"))
source(file.path(TOP, "sars2vl", "data_utils.R"))

param_sizes_all_zero_current <- TRUE
param_sizes_suffix <- ifelse(param_sizes_all_zero_current, "", "GroupDiff")
plot_dir_final <- ifelse(param_sizes_all_zero_current, "param_sizes_all_zero",
  "param_sizes_not_zero"
)
plot_dir <- file.path(DIR_FIGURES_VL_SIM, plot_dir_final)

# Plot params
alpha_current <- ALPHA_SIM_PLOT
width_current <- WIDTH_JITTER_SIM
height_current <- HEIGHT_JITTER_SIM
size_current <- GEOM_POINT_SIZE_SIM
height_plot <- HEIGHT_SIM_PLOT
left_margin <- ANNOTATION_MARGIN_LEFT_SIM_PLOT
hjust <- ANNOTATION_HJUST_SIM_PLOT
dodge_height <- 0.5
col_width <- COL_WIDTH
```

# Utilities
```{r}
# Function to format a single row as LaTeX with dec_places decimal digits for numeric values
format_row <- function(row, bold = FALSE, dec_places = 2, latex_math = FALSE) {
  format_string <- paste0("%.", dec_places, "f")
  # Format numeric values to dec_places decimal places
  formatted_row <- sapply(row, function(x) {
    if (is.numeric(x)) {
      # Format numeric values to dec_places decimal places, $s are necessary to
      # format minus signs correctly in latex
      x <- sprintf(format_string, x)
    }
    if (bold) {
      assert_that(!latex_math)
      x <- paste0("\\textbf{", x, "}")
    } else {
      x <- x
    }
    if (latex_math) {
      x <- paste0("$", x, "$")
    }
    return(x)
  })
  # Collapse the row with " & " and add \\ \hline
  paste0(paste0(formatted_row, collapse = " & "), " \\\\ \\hline")
}
```



# Load data

```{r}
model_no_current <- 2

dt_orig <- process_sim_dt(
  dt = fread(PARAMS_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dte_orig <- process_sim_dt(
  dt = fread(PARAMS_EXP_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dthdi_orig <- process_sim_dt(
  dt = fread(PARAMS_HDI_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)

dt_all <- process_sim_dt(
  dt = fread(PARAMS_DIFF_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dtr_all <- process_sim_dt(fread(PARAMS_REL_DIFF_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dthdi_all <- process_sim_dt(fread(PARAMS_DIFF_HDI_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dthdi50_all <- process_sim_dt(fread(PARAMS_DIFF_HDI50_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dtrhdi_all <- process_sim_dt(fread(PARAMS_REL_DIFF_HDI_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)
dtrhdi50_all <- process_sim_dt(fread(PARAMS_REL_DIFF_HDI50_FILE, sep = "\t"),
  param_sizes_all_zero = param_sizes_all_zero_current
)

dtr_all <- merge(dtr_all, dtrhdi_all, by = "model", suffix = c("", "_hdi"))
dtr_all <- parse_hdi_cols(merge(dtr_all, dtrhdi50_all, by = "model", suffix = c("", "_hdi50")))

dt_all <- merge(dt_all, dthdi_all, by = "model", suffix = c("", "_hdi"))
dt_all <- parse_hdi_cols(merge(dt_all, dthdi50_all, by = "model", suffix = c("", "_hdi50")))

dts_model <- get_sim_diff_dts(
  model_no = model_no_current,
  param_sizes_all_zero = param_sizes_all_zero_current
)
dt <- dts_model$dt
dtr <- dts_model$dtr


annotation_label <- ifelse(model_no_current %in% c(3, 1), "A", "B")
```

# Plots

## Differences between true and estimated main parameter values

```{r, fig.width=col_width*2, fig.height=height_plot}
p_up <- plot_main_model_params(dtr,
  y = "min_n_tests", x = "time2peak",
  color = "min_n_pos_tests", alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL2,
  xlabel = "Relative error in proliferation time (%)",
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT,
  as_percent = TRUE
) +
  xlim(0, 55)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simTimeToPeakDiffModel",
    model_no_current,
    param_sizes_suffix, ".png"
  )),
  plot = p_up,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_up
```



```{r, fig.width=COL_WIDTH*2, fig.height=height_plot}
p_down <- plot_main_model_params(dtr,
  y = "min_n_tests", x = "time_from_peak",
  color = "min_n_pos_tests", alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = "Relative error in clearance time (%)",
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT,
  as_percent = TRUE
) + xlim(5, 30)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simTimeFromPeakDiffModel",
    model_no_current,
    param_sizes_suffix,
    ".png"
  )),
  plot = p_down,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_down
```


```{r, fig.width=COL_WIDTH*2, fig.height=height_plot}
xlabel_peak <- expression(atop(
  "Estimated - true peak ",
  phantom(x)^phantom(x) * paste(log[10], " viral load")
))
p_peak <- plot_main_model_params(dt,
  y = "min_n_tests", x = "intercept",
  color = "min_n_pos_tests", alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = xlabel_peak,
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT
)

ggsave(
  filename = file.path(plot_dir, paste0(
    "simPeakDiffModel",
    model_no_current,
    param_sizes_suffix, ".png"
  )),
  plot = p_peak,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_peak
```

## Differences between true and estimated sigma parameter values

```{r}
xlabel_up_sigma <- "Relative difference in random up-slope\nstandard deviation"
p_up_sigma <- plot_main_model_params(dtr,
  y = "min_n_tests", 
  x = "slope_up_sigma",
  color = "min_n_pos_tests", 
  alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = xlabel_up_sigma,
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT
)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simUpSlopeSigmaDiffModel",
    model_no_current, ".png"
  )),
  plot = p_up_sigma,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_up_sigma
```

```{r}
xlabel_down_sigma <- "Relative difference in random down-slope\nstandard deviation"
p_down_sigma <- plot_main_model_params(dtr,
  y = "min_n_tests", x = "slope_down_sigma",
  color = "min_n_pos_tests", alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = xlabel_down_sigma,
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT
)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simDownSlopeSigmaDiffModel",
    model_no_current, ".png"
  )),
  plot = p_down_sigma,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_down_sigma
```


```{r}
xlabel_peak_sigma <- "Relative difference in random intercept\nstandard deviation"
p_peak_sigma <- plot_main_model_params(dtr,
  y = "min_n_tests", x = "intercept_sigma",
  color = "min_n_pos_tests", alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = xlabel_peak_sigma,
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT
)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simPeakSigmaDiffModel",
    model_no_current, ".png"
  )),
  plot = p_peak_sigma,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_peak_sigma
```

```{r}
xlabel_sigma <- "Relative difference in residual error"
p_sigma <- plot_main_model_params(dt,
  y = "min_n_tests", 
  x = "sigma",
  color = "min_n_pos_tests", 
  alpha = alpha_current,
  height_dodge = dodge_height,
  size_point = size_current,
  ylabel = MINIMUM_PCRS_LABEL,
  xlabel = xlabel_sigma,
  color_label = MINIMUM_POS_PCRS_LABEL2,
  annotation_label = annotation_label,
  left_margin = ANNOTATION_MARGIN_LEFT_SIM_PLOT,
  hjust = ANNOTATION_HJUST_SIM_PLOT
)


ggsave(
  filename = file.path(plot_dir, paste0(
    "simSigmaDiffModel",
    model_no_current, ".png"
  )),
  plot = p_sigma,
  width = COL_WIDTH * 2,
  height = height_plot,
  units = "in",
  dpi = DPI
)
```

```{r}
p_sigma
```


## Differences between true and estimated group params

```{r}
group_var <- "omicron"
group_label <- "omicron"
up_var <- paste0("slope_up_", group_var)
down_var <- paste0("slope_down_", group_var)
peak_var <- paste0("intercept_", group_var)

up_var2 <- paste0("slope_up_", group_var, "_hdi_contains_true_value")
down_var2 <- paste0("slope_down_", group_var, "_hdi_contains_true_value")
peak_var2 <- paste0("intercept_", group_var, "_hdi_contains_true_value")

shapes <- c(16, 15)
```


```{r}
p_up_group <- ggplot(dtr, aes(
  x = min_n_tests, y = !!sym(up_var),
  color = min_n_pos_tests, shape = !!sym(up_var2)
)) +
  scale_color_viridis_d() + # Default: viridis palette
  geom_jitter(width = width_current, height = height_current, alpha = alpha_current, size = size_current) +
  labs(
    title = "", x = MINIMUM_PCRS_LABEL,
    y = paste0("Relative difference in ", group_label, "\nup-slope parameter"),
    color = MINIMUM_POS_PCRS_LABEL2,
    shape = "94% HDI contains\ntrue value"
  ) +
  scale_shape_manual(values = rev(shapes)) +
  theme_thesis()


ggsave(
  filename = file.path(plot_dir, paste0(
    "simUpSlope", toTitleCase(group_var),
    "DiffModel", model_no_current, ".png"
  )),
  plot = p_up_group,
  width = COL_WIDTH * 2,
  height = height_plot * 1.5,
  units = "in",
  dpi = DPI
)
```

```{r}
p_up_group
```


```{r}
p_down_group <- ggplot(dtr, aes(
  x = min_n_tests, y = !!sym(down_var),
  color = min_n_pos_tests
)) +
  scale_color_viridis_d() + # Default: viridis palette
  geom_jitter(width = width_current, height = height_current, alpha = alpha_current, size = size_current) +
  labs(
    title = "", x = MINIMUM_PCRS_LABEL,
    y = paste0("Relative difference in ", group_label, "\ndown-slope parameter"),
    color = MINIMUM_POS_PCRS_LABEL2
  ) +
  theme_thesis()


ggsave(
  filename = file.path(plot_dir, paste0(
    "simDownSlope", toTitleCase(group_var), "PAMSDiffModel",
    model_no_current, ".png"
  )),
  plot = p_down_group,
  width = COL_WIDTH * 2,
  height = height_plot * 1.5,
  units = "in",
  dpi = DPI
)
```

```{r}
p_down_group
```

```{r}
p_peak_group <- ggplot(dtr, aes(
  x = min_n_tests, y = !!sym(peak_var),
  color = min_n_pos_tests
)) +
  scale_color_viridis_d() + # Default: viridis palette
  geom_jitter(width = width_current, height = height_current, alpha = alpha_current, size = size_current) +
  labs(
    title = "", x = MINIMUM_PCRS_LABEL,
    y = paste0("Relative difference in ", group_label, "\nintercept parameter"),
    color = MINIMUM_POS_PCRS_LABEL2
  ) +
  theme_thesis()


ggsave(
  filename = file.path(plot_dir, paste0(
    "simIntercept", toTitleCase(group_var),
    "DiffModel", model_no_current, ".png"
  )),
  plot = p_peak_group,
  width = COL_WIDTH * 2,
  height = height_plot * 1.5,
  units = "in",
  dpi = DPI
)
```

```{r}
p_peak_group
```

### Summary plots

```{r}
generate_sim_data_group_plots(param_sizes_all_zero = TRUE)
```

```{r}
p_group_sum
```

```{r}
p_slope_up_box
```


# Tables

```{r}
cols_of_interest <- c(
  "model_no_stan", "min_n_tests", "min_n_pos_tests", "iteration",
  "intercept", "intercept_hdi", "time2peak", "time2peak_hdi",
  "time_from_peak", "time_from_peak_hdi", "sigma"
)

cols_of_interest2 <- c(
  "model_no_stan", "min_n_tests", "min_n_pos_tests", "iteration",
  "intercept", "time2peak", "time_from_peak", "sigma"
)

dt_table <- dtr_all[param_sizes_all_zero == param_sizes_all_zero_current]
dt_table <- dt_table[, ..cols_of_interest]
dt_table[, model_no_stan_str := as.character(model_no_stan)]

# Rounding
dt_table[, c("intercept", "time2peak", "time_from_peak") := lapply(.SD, round, 3),
  .SDcols = c("intercept", "time2peak", "time_from_peak")
]
dt_table[, c("sigma") := lapply(.SD, round, 3),
  .SDcols = c("sigma")
]

# Multiplying
dt_table[, c("intercept", "time2peak", "time_from_peak") := lapply(.SD, function(x) sprintf("%.1f", x * 100)),
  .SDcols = c("intercept", "time2peak", "time_from_peak")
]
dt_table[, c("sigma") := lapply(.SD, function(x) x * 100),
  .SDcols = c("sigma")
]

for (col_current in c("intercept", "time2peak", "time_from_peak")) {
  col_hdi_current <- paste0(col_current, "_hdi")
  dt_table[, lower := round(as.numeric(sapply(get(col_hdi_current), `[`, 1)), 3) * 100] # First element
  dt_table[, upper := round(as.numeric(sapply(get(col_hdi_current), `[`, 2)), 3) * 100] # Second element
  # Create new column containing median and hpdi
  dt_table[, (col_current) := paste0(
    "$", get(col_current), "$",
    " & ",
    "$", sprintf("%.1f", lower), "$",
    " & ",
    "$", sprintf("%.1f", upper), "$"
  )]
}


dt_table <- dt_table[, ..cols_of_interest2]

replace_colnames <- list("model_no_stan_str" = "Model", "min_n_tests" = "Min. PCR days", "min_n_pos_tests" = "Min. pos. PCR days", "iteration" = "Iteration", "intercept" = "Peak", "time2peak" = "Days to peak", "time_from_peak" = "Days from peak", "slope_up_sigma" = "Random up-slope standard deviation", "slope_down_sigma" = "Random down-slope standard deviation", "intercept_sigma" = "Random intercept standard deviation", "sigma" = "Residual standard deviation")

# Replace model names
dt_table[model_no_stan == 1, model_no_stan_str := "1"]
dt_table[model_no_stan == 3, model_no_stan_str := "1_2"]
dt_table[model_no_stan == 2, model_no_stan_str := "2"]
dt_table <- dt_table[, c(9, 1, 2, 3, 4, 5, 6, 7, 8)]
dt_table <- dt_table[order(model_no_stan_str, iteration, min_n_tests, min_n_pos_tests)]


# Replace column names
setnames(dt_table,
  old = names(replace_colnames), new = unlist(replace_colnames),
  skip_absent = TRUE
)

if (param_sizes_all_zero_current == FALSE) {
  dt_table[, Iteration := NULL]
}
```

## Differences between true and estimated main parameter values

```{r}
if (param_sizes_all_zero_current) {
  model_params_current_latex <- c(
    "Min. PCR days", "Min. pos. PCR days",
    "Peak", "Days to peak", "Days from peak"
  )
} else {
  model_params_current_latex <- c(
    "Min. PCR days", "Min. pos. PCR days",
    "Peak", "Days to peak", "Days from peak"
  )
}

model_no_current_latex <- c("1", "2")
model_no_current_latex_label <- ifelse(model_no_current_latex == "1_2", "1_2", "1-2")
dt_table_current <- dt_table[Model %in% model_no_current_latex]
dt_table_current <- dt_table_current[, ..model_params_current_latex]

param_sizes_all_zero_suffix <- ifelse(param_sizes_all_zero_current, "", "GroupDiffs")
label_current <- paste0("simulationEstimation", param_sizes_all_zero_suffix, model_no_current_latex_label)


# Split the data by the "group" column
grouped_data <- split(dt_table_current, dt_table_current[["Min. PCR days"]])

# Generate the LaTeX body for each group and combine with thick horizontal lines
latex_body <- paste(
  lapply(grouped_data, function(group_df) {
    # Apply format_row to each row in the group and collapse with "\n"
    paste(apply(group_df, 1, format_row), collapse = "\n")
  }),
  collapse = " \n\\specialrule{0.55pt}{0pt}{0pt}\n"
)
```

### Median deviations

```{r}
# Compute median deviations for different models

relative_difference <- TRUE
if (relative_difference) {
  dt_current <- dtr_all
  results_factor <- 100
} else {
  dt_current <- dt_all
  results_factor <- 1
}
dt_current_stats <- dt_current[param_sizes_all_zero == param_sizes_all_zero_current]
dt_current_stats[, n_tests_combination := paste0(min_n_tests, "_", min_n_pos_tests)]
by_var_current <- "model_no_stan" # "n_tests_combination"#


result_peak <- dt_current_stats[
  , .(
    median_peak_dev = median(intercept) * eval(results_factor),
    lower_quartile = quantile(intercept, 0.25, na.rm = TRUE) * eval(results_factor),
    upper_quartile = quantile(intercept, 0.75, na.rm = TRUE) * eval(results_factor),
    iqr = IQR(intercept * eval(results_factor), na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]
result_time2peak <- dt_current_stats[
  , .(
    median_time2peak_dev = median(time2peak) * eval(results_factor),
    lower_quartile = quantile(time2peak, 0.25, na.rm = TRUE) * eval(results_factor),
    upper_quartile = quantile(time2peak, 0.75, na.rm = TRUE) * eval(results_factor),
    iqr = IQR(time2peak * eval(results_factor), na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]
result_time_from_peak <- dt_current_stats[
  , .(
    median_time_from_peak_dev = median(time_from_peak) * eval(results_factor),
    lower_quartile = quantile(time_from_peak, 0.25, na.rm = TRUE) * eval(results_factor),
    upper_quartile = quantile(time_from_peak, 0.75, na.rm = TRUE) * eval(results_factor),
    iqr = IQR(time_from_peak * eval(results_factor), na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]
```


```{r}
result_time2peak
```

#### Between iterations of model 2

```{r}
# Evaluate variation between different iterations (was done for Model 2 only)
relative_difference <- TRUE
if (relative_difference) {
  dt_current <- dtr_all
  results_factor <- 100
} else {
  dt_current <- dt_all
  results_factor <- 1
}
dt_current_stats_2 <- dt_current[(param_sizes_all_zero == param_sizes_all_zero_current) &
  (model_no_stan == 2)]


diff_2 <- dt_current_stats_2[,
  {
    peak1 <- intercept[iteration == 1] * eval(results_factor)
    peak2 <- intercept[iteration == 2] * eval(results_factor)
    diff_peak <- peak1 - peak2
    time2_1 <- time2peak[iteration == 1] * eval(results_factor)
    time2_2 <- time2peak[iteration == 2] * eval(results_factor)
    diff_time2 <- time2_1 - time2_2
    time_from_1 <- time_from_peak[iteration == 1] * eval(results_factor)
    time_from_2 <- time_from_peak[iteration == 2] * eval(results_factor)
    diff_time_from <- time_from_1 - time_from_2
    list(diff_peak = diff_peak, diff_time2 = diff_time2, diff_time_from = diff_time_from)
  },
  by = .(min_n_tests, min_n_pos_tests)
]

result_2 <- diff_2[
  , .(
    median_peak_dev = median(diff_peak),
    lower_quartile_peak = quantile(diff_peak, 0.25, na.rm = TRUE),
    upper_quartile_peak = quantile(diff_peak, 0.75, na.rm = TRUE),
    iqr_peak = IQR(diff_peak, na.rm = TRUE),
    median_time2_dev = median(diff_time2),
    lower_quartile_time2 = quantile(diff_time2, 0.25, na.rm = TRUE),
    upper_quartile_time2 = quantile(diff_time2, 0.75, na.rm = TRUE),
    iqr_time2 = IQR(diff_time2, na.rm = TRUE),
    median_time_from_dev = median(diff_time_from),
    lower_quartile_time_from = quantile(diff_time_from, 0.25, na.rm = TRUE),
    upper_quartile_time_from = quantile(diff_time_from, 0.75, na.rm = TRUE),
    iqr_time_from = IQR(diff_time_from, na.rm = TRUE)
  )
]
```

```{r}
result_2
```


## Differences between true and estimated sigma parameter values

```{r}
cols_of_interest <- c(
  "model_no_stan", "min_n_tests", "min_n_pos_tests", "iteration",
  "intercept_sigma", "slope_up_sigma", "slope_down_sigma", "sigma"
)

dtr_table <- dtr_all[param_sizes_all_zero == param_sizes_all_zero_current]
dtr_table <- dtr_table[, ..cols_of_interest]

replace_colnames <- list("model_no_stan_str" = "Model", "min_n_tests" = "Min. PCR days", "min_n_pos_tests" = "Min. pos. PCR days", "iteration" = "Iteration", "intercept" = "Peak", "time2peak" = "Days to peak", "time_from_peak" = "Days from peak", "slope_up_sigma" = "Random up-slope standard deviation", "slope_down_sigma" = "Random down-slope standard deviation", "intercept_sigma" = "Random intercept standard deviation", "sigma" = "Residual standard deviation")

# Replace model names
dtr_table[model_no_stan == 3, model_no_stan_str := "1_2"]
dtr_table[model_no_stan == 2, model_no_stan_str := "2"]
dtr_table[model_no_stan == 1, model_no_stan_str := "1"]
dtr_table <- dtr_table[, c(9, 1, 2, 3, 4, 5, 6, 7, 8)]
dtr_table <- dtr_table[order(model_no_stan_str, iteration, min_n_tests, min_n_pos_tests)]

# Rounding
dtr_table[, c("intercept_sigma", "slope_up_sigma", "slope_down_sigma", "sigma") := lapply(.SD, round, 3),
  .SDcols = c("intercept_sigma", "slope_up_sigma", "slope_down_sigma", "sigma")
]
dtr_table[, c("intercept_sigma", "slope_up_sigma", "slope_down_sigma", "sigma") := lapply(
  .SD,
  function(x) x * 100
),
.SDcols = c("intercept_sigma", "slope_up_sigma", "slope_down_sigma", "sigma")
]

# Replace column names
setnames(dtr_table, old = names(replace_colnames), new = unlist(replace_colnames), skip_absent = TRUE)

if (param_sizes_all_zero_current) {
  model_params_current_latex <- c(
    "Model", "Min. PCR days", "Min. pos. PCR days",
    "Iteration", "Random up-slope standard deviation",
    "Random down-slope standard deviation",
    "Random intercept standard deviation",
    "Residual standard deviation"
  )
} else {
  model_params_current_latex <- c(
    "Model", "Min. PCR days", "Min. pos. PCR days",
    "Random up-slope standard deviation",
    "Random down-slope standard deviation",
    "Random intercept standard deviation",
    "Residual standard deviation"
  )
}


dtr_table_current <- dtr_table[, ..model_params_current_latex]
param_sizes_all_zero_suffix <- ifelse(param_sizes_all_zero_current, "", "GroupDiffs")
label_current <- paste0("simulationEstimationSigma", param_sizes_all_zero_suffix)


# Split the data by the "group" column
grouped_data <- split(dtr_table_current, dtr_table_current[["Min. PCR days"]])

# Generate the LaTeX body for each group and combine with thick horizontal lines
latex_body <- paste(
  lapply(grouped_data, function(group_df) {
    # Apply format_row to each row in the group and collapse with "\n"
    paste(apply(group_df, 1, function(row) {
      format_row(row,
        bold = FALSE, dec_places = 3,
        latex_math = TRUE
      )
    }), collapse = "\n")
  }),
  collapse = " \n\\specialrule{0.55pt}{0pt}{0pt}\n"
)
```

### Median deviations

```{r}
# Compute median deviations for different models

dt_current_stats <- dtr_all[param_sizes_all_zero == param_sizes_all_zero_current]
dt_current_stats[, n_tests_combination := paste0(min_n_tests, "_", min_n_pos_tests)]
by_var_current <- "model_no_stan"

result_peak_sigma <- dt_current_stats[
  , .(
    median_dev = median(intercept_sigma),
    lower_quartile = quantile(intercept_sigma, 0.25, na.rm = TRUE),
    upper_quartile = quantile(intercept_sigma, 0.75, na.rm = TRUE),
    iqr = IQR(intercept_sigma, na.rm = TRUE),
    max = max(intercept_sigma),
    min = min(intercept_sigma)
  ),
  by = eval(as.name(by_var_current))
]
result_slope_up_sigma <- dt_current_stats[
  , .(
    median_dev = median(slope_up_sigma),
    lower_quartile = quantile(slope_up_sigma, 0.25, na.rm = TRUE),
    upper_quartile = quantile(slope_up_sigma, 0.75, na.rm = TRUE),
    iqr = IQR(slope_up_sigma, na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]
result_slope_down_sigma <- dt_current_stats[
  , .(
    median_dev = median(slope_down_sigma),
    lower_quartile = quantile(slope_down_sigma, 0.25, na.rm = TRUE),
    upper_quartile = quantile(slope_down_sigma, 0.75, na.rm = TRUE),
    iqr = IQR(slope_down_sigma, na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]

result_sigma <- dt_current_stats[
  , .(
    median_dev = median(sigma),
    lower_quartile = quantile(sigma, 0.25, na.rm = TRUE),
    upper_quartile = quantile(sigma, 0.75, na.rm = TRUE),
    iqr = IQR(sigma, na.rm = TRUE)
  ),
  by = eval(as.name(by_var_current))
]
```
 
```{r}
result_sigma
```
 
 
# Group-specific parameter estimates

## Proportion of wrongly estimated differences with HDI excluding 0

```{r}
exclude_true_zero_values <- TRUE
column_suffix <- ifelse(param_sizes_all_zero_current,
  "_hdi_contains_true_value",
  "_hdi_combination"
)
group_cols_hdi_mapping <- return_group_cols_hdi_mapping(
  log_prefix = "", include_base_cols = FALSE,
  column_suffix = column_suffix
)
group_cols_intercept <- return_group_cols("")$group_cols_mapping$intercept
group_cols_intercept <- c(
  group_cols_intercept, paste0(group_cols_intercept, "_hdi"),
  paste0(group_cols_intercept, "_hdi50")
)
group_cols_slope_up <- return_group_cols("")$group_cols_mapping$slope_up
group_cols_slope_up <- c(
  group_cols_slope_up, paste0(group_cols_slope_up, "_hdi"),
  paste0(group_cols_slope_up, "_hdi50")
)
group_cols_slope_down <- return_group_cols("")$group_cols_mapping$slope_down
group_cols_slope_down <- c(
  group_cols_slope_down, paste0(group_cols_slope_down, "_hdi"),
  paste0(group_cols_slope_down, "_hdi50")
)


dts_current_stats <- get_sim_dts(
  model_no = NULL,
  param_sizes_all_zero = param_sizes_all_zero_current
)
dtr_current_stats <- dts_current_stats$dte

if (param_sizes_all_zero_current) {
  combination_cols <- FALSE
  contains_true_value_cols <- TRUE
  hdi_col_name <- "hdi_contains_true_value"
} else {
  combination_cols <- TRUE
  contains_true_value_cols <- FALSE
  hdi_col_name <- "hdi_combination"
}

dtr_intercept <- dtr_current_stats[, c(
  "model_no_stan", group_cols_hdi_mapping$intercept,
  group_cols_intercept
), with = FALSE]
dtr_intercept <- reshape_dt(dtr_intercept,
  id_vars = c("model_no_stan"),
  combination_cols = combination_cols,
  contains_true_value_cols = contains_true_value_cols
)
dtr_intercept <- dtr_intercept[, c("model_no_stan", hdi_col_name, "statistic", "value"),
  with = FALSE
]


dtr_slope_up <- dtr_current_stats[, c(
  "model_no_stan", group_cols_hdi_mapping$slope_up,
  group_cols_slope_up
), with = FALSE]
dtr_slope_up <- reshape_dt(dtr_slope_up,
  id_vars = c("model_no_stan"),
  combination_cols = combination_cols,
  contains_true_value_cols = contains_true_value_cols
)
dtr_slope_up <- dtr_slope_up[, c("model_no_stan", hdi_col_name, "statistic", "value"), with = FALSE]


dtr_slope_down <- dtr_current_stats[, c(
  "model_no_stan", group_cols_hdi_mapping$slope_down,
  group_cols_slope_down
), with = FALSE]
dtr_slope_down <- reshape_dt(dtr_slope_down,
  id_vars = c("model_no_stan"),
  combination_cols = combination_cols,
  contains_true_value_cols = contains_true_value_cols
)
dtr_slope_down <- dtr_slope_down[, c("model_no_stan", hdi_col_name, "statistic", "value"),
  with = FALSE
]


if (param_sizes_all_zero_current) {
  by_stat <- "model_no_stan"
} else {
  by_stat <- c("model_no_stan", "hdi_combination")
}

if (param_sizes_all_zero_current) {
  result_intercept <- dtr_intercept[
    , .(
      mean = mean(hdi_contains_true_value)
    ),
    by = model_no_stan
  ]

  result_slope_up <- dtr_slope_up[
    , .(
      mean = mean(hdi_contains_true_value)
    ),
    by = model_no_stan
  ]

  result_slope_down <- dtr_slope_down[
    , .(
      mean = mean(hdi_contains_true_value)
    ),
    by = model_no_stan
  ]
} else {
  # Compute the relative frequency of the values present in combination columns
  interaction_vars <- c("hospitalized_alpha", "hospitalized_delta", "hospitalized_omicron")
  intercept_interaction_vars <- paste0("intercept_", interaction_vars)
  slope_up_interaction_vars <- paste0("slope_up_", interaction_vars)
  slope_down_interaction_vars <- paste0("slope_down", interaction_vars)
  # Step 1: Count occurrences
  if (exclude_true_zero_values) {
    # Exclude Alpha and Unknown and interaction parameters as their true values
    # are 0 and we want the percentage of DIFFERENCES that were correctly identified as such
    counts_intercept <- dtr_intercept[
      !statistic %in% c(
        c(
          "intercept_alpha",
          "intercept_unknown",
          "intercept_gender"
        ),
        intercept_interaction_vars
      ),
      .(count = .N),
      by = .(model_no_stan, hdi_combination)
    ]
  } else {
    # For the percentage of cases where the estimate excluded both the true value and 0,
    # we use all parameters again
    counts_intercept <- dtr_intercept[, .(count = .N),
      by = .(model_no_stan, hdi_combination)
    ]
  }

  # # Step 2: Compute relative frequencies
  counts_intercept[, relative_freq := count / sum(count), by = model_no_stan]
  counts_intercept <- counts_intercept[order(model_no_stan, hdi_combination)]

  # Step 1: Count occurrences
  if (exclude_true_zero_values) {
    counts_slope_up <- dtr_slope_up[
    !statistic %in% c(
      c(
        "slope_up_pams",
        "slope_up_hospitalized",
        "slope_up_age_cat2",
        "slope_up_age_cat3",
        "slope_up_gender"
      ),
      slope_up_interaction_vars
    ),
    .(count = .N),
    by = .(model_no_stan, hdi_combination)
    ]
  } else {
    counts_slope_up <- dtr_slope_up[, .(count = .N),
      by = .(model_no_stan, hdi_combination)
      ]
  }

  # # Step 2: Compute relative frequencies
  counts_slope_up[, relative_freq := count / sum(count), by = model_no_stan]
  counts_slope_up <- counts_slope_up[order(model_no_stan, hdi_combination)]

  # Step 1: Count occurrences
  if (exclude_true_zero_values) {
    counts_slope_down <- dtr_slope_down[
    !statistic %in% c(
      c("slope_down_gender"),
      slope_down_interaction_vars
    ),
    .(count = .N),
    by = .(model_no_stan, hdi_combination)
    ]
  } else {
    counts_slope_down <- dtr_slope_down[, .(count = .N),
      by = .(model_no_stan, hdi_combination)
      ]
  }

  # Step 2: Compute relative frequencies
  counts_slope_down[, relative_freq := count / sum(count), by = model_no_stan]
  counts_slope_down <- counts_slope_down[order(model_no_stan, hdi_combination)]
}
```

```{r}
counts_slope_up
```


## Median (absolute) deviations for group-specific params (by parameter type)

```{r}
# Compute median deviations for different models

group_cols <- return_group_cols("")
id_var_current <- "model_no_stan" # "n_tests_combination"

dtr_current_stats <- dtr_all[param_sizes_all_zero == param_sizes_all_zero_current]
dtr_current_stats[, n_tests_combination := paste0(min_n_tests, "_", min_n_pos_tests)]

group_cols_current <- c(
  id_var_current,
  group_cols$group_cols_mapping$intercept,
  paste0(group_cols$group_cols_mapping$intercept, "_hdi"),
  paste0(group_cols$group_cols_mapping$intercept, "_hdi50")
)
dtr_intercept <- dtr_current_stats[, group_cols_current, with = FALSE]
dtr_intercept <- reshape_dt(dtr_intercept, id_vars = id_var_current, hdi_cols = TRUE)

group_cols_current <- c(
  id_var_current, group_cols$group_cols_mapping$slope_up,
  paste0(group_cols$group_cols_mapping$slope_up, "_hdi"),
  paste0(group_cols$group_cols_mapping$slope_up, "_hdi50")
)
dtr_slope_up <- dtr_current_stats[, group_cols_current, with = FALSE]
dtr_slope_up <- reshape_dt(dtr_slope_up, id_vars = id_var_current, hdi_cols = TRUE)

group_cols_current <- c(
  id_var_current, group_cols$group_cols_mapping$slope_down,
  paste0(group_cols$group_cols_mapping$slope_down, "_hdi"),
  paste0(group_cols$group_cols_mapping$slope_down, "_hdi50")
)
dtr_slope_down <- dtr_current_stats[, group_cols_current, with = FALSE]
dtr_slope_down <- reshape_dt(dtr_slope_down,
  id_vars = id_var_current,
  hdi_cols = TRUE
)


result_intercept <- dtr_intercept[
  , .(
    median = median(abs(value))
  ),
  by = eval(as.name(id_var_current))
] %>%
  .[order(median)]

result_slope_up <- dtr_slope_up[
  , .(
    median = median(abs(value))
  ),
  by = eval(as.name(id_var_current))
] %>%
  .[order(median)]

result_slope_down <- dtr_slope_down[
  , .(
    median = median(abs(value))
  ),
  by = eval(as.name(id_var_current))
] %>%
  .[order(median)]
```

```{r}
result_slope_down
```

## Correlations of interaction effects

```{r}
# Correlations between variant estimates, hospitalization estimates and hospitalization:variant estimates

param_var <- "intercept"
variant_var <- "omicron"


dt_cor <- data.table(variant = character(), param = character(), cor = numeric())

for (param_var in c("intercept", "slope_up", "slope_down")) {
  for (variant_var in c("alpha", "delta", "omicron")) {
    variant_param <- paste0(param_var, "_", variant_var)
    variant_hosp_param <- paste0(param_var, "_hospitalized_", variant_var)
    cor_now <- cor(dte_orig[[variant_param]], dte_orig[[variant_hosp_param]])
    # New row(s) to add (as a data.table or list)
    new_row <- data.table(variant = variant_var, param = param_var, cor = cor_now)
    # Append the new row
    dt_cor <- rbind(dt_cor, new_row, fill = TRUE)
  }
}

dt_cor[, mean_cor_by_param := mean(cor), by = .(param)]
dt_cor[, mean_cor_by_variant := mean(cor), by = .(variant)]
```

